Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}

### 2016-12-16 ident_manu.Rnw
elsarticle class

\documentclass[letterpaper,12pt,oneside]{article}
\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{indentfirst}
\usepackage{titlesec}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{outlines}
\usepackage{lineno}
\usepackage{array}
\usepackage{times}
\usepackage{cleveref}
\usepackage{acronym}
\usepackage[position=t]{subfig}
\usepackage{paralist}
\usepackage[noae]{Sweave}
\usepackage{natbib}
\usepackage{array}
\usepackage{pdflscape}
\usepackage{bm}
\usepackage{fixltx2e}
% \usepackage[inline]{showlabels}
\bibpunct{(}{)}{,}{a}{}{,}

% page margins and section title formatting
\linespread{1.5}
\setlength{\footskip}{0.5in}
\titleformat*{\section}{\Large\bf\em}
\titleformat*{\subsection}{\singlespace\large\bf}
\titleformat*{\subsubsection}{\singlespace\normalsize\bf\em}
\titlespacing{\section}{0in}{0in}{0in}
\titlespacing{\subsection}{0in}{0in}{0in}
\titlespacing{\subsubsection}{0in}{0in}{0in}

@@

@@

% get the version based on commit date
<<echo = FALSE, cache = FALSE>>=
raw <- system('git log -1', intern = TRUE)
raw <- raw[grep('^Date', raw)]
raw <- paste('Version', raw)
@

@@

\raggedbottom
\linenumbers
\raggedright
\urlstyle{same}
\setlength{\parindent}{0.5in}
\renewcommand\refname{References \vspace{12pt}}

\begin{singlespace}
\title{{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, John C. Lehrter$^2$, Lisa L. Lowe$^3$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Email: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}}}
	  \\\\{\textit {\normalsize $^2$Dauphin Island Sea Lab, University of South Alabama}}
  \\{\textit {\normalsize Dauphin Island, AL 36528}}
	\\{\textit {\normalsize Phone: 251-861-2141 ext. 7567}}
	\\{\textit {\normalsize Email: \href{mailto:jlehrter@disl.org}{jlehrter@disl.org}}}
	\\\\{\textit {\normalsize $^3$Lockheed Martin IS \& GS - Civil, supporting the USEPA}}
	\\{\textit {\normalsize Resarch Triangle Park, NC 27709}}
	\\{\textit {\normalsize Phone: 919-541-3985,}}
	\\{\textit {\normalsize Email: \href{mailto:lowe.lisa@epa.gov}{lowe.lisa@epa.gov}}}
  \vspace{1in} 
  \\ \Sexpr{raw}
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

\begin{abstract}
\noindent Bio-geo-chemical models are useful tools in environmental sciences that can guide management and policy-making. Consequently, significant time and resources are spent developing these models in system-specific contexts. The optimization of model parameters to maximize precision, including transferability of these models to different systems, are fundamental concerns in the development and application of these tools. This study provides a context for understanding quantitative limitations of coupled hydrodynamic-ecological models by evaluating parameter sensitivity and identifability of a \ac{zerod} unit of a larger spatio-temporal model of hypoxia on the \acl{lcs} of Gulf of Mexico.  The analysis provides a contrast of numeric and ecological certainty in parameter subsets using a systematic framework to infer larger trends in dissolved oxygen dynamics over time, having implications for understanding factors that contribute to environmental conditions that are detrimental to aquatic resources.  In particular, we focus on issues of parameter identifiability using local sensitivity analyses to provide quantitative descriptions of numerical constraints on model precision.  The sensitivity of state variables differed considerably with parameter changes, although most variables were responsive to changes in parameters that influenced planktonic growth rates.  Variation in sensitivity had a direct correspendence with identifiability, such that only small subsets of the complete parameter set were characterized as having unique effects on the model output. As a result, we provide a set of parameter selection heuristics that can be used to identify parameters for model calibration that depend on relative sensitivity and ecological categories within the biogeochemical equations. Although these concerns have been expressed in the literature, they are rarely explicitly addressed or included in evaluations of water quality models.  In addition to immediate implications for regional models, we provide a framework for describing the effects of parameter uncertainty and identifiability that can be applied to similar models to better inform environmental management.
\end{abstract}
Keywords: \ac{cgem}, \ac{gom}, Hypoxia, Identifiability, Sensitivity
\acresetall

@@

\clearpage
\begin{singlespace}
\bibliographystyle{apalike_mine}
\bibliography{refs}
\end{singlespace}

@@

% parameter exclusion temp
<<exclex, eval = T, fig.cap = 'Identifiability (as $\\gamma$, \\cref{gameq}) for \\ac{do} of organic matter parameters for subset combinations in \\cref{fig:identplo}.  Identifiability is evaluated for subsets that excluded and included the parameters at the top of each plot. Identifiability of including all eight parameters is in \\cref{fig:identplo}. Grey lines indicate potential thresholds at $\\gamma = 10, 15$ for maximum acceptable identifiability.', fig.height = 7.25, fig.width = 7.25, echo = F, cache = TRUE, out.width = "\\textwidth">>=
data(cat_ident_all)

>>>>

\documentclass[preprint]{elsarticle}
\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
\usepackage{lineno}
\usepackage{cleveref}
\usepackage{acronym}
\usepackage{paralist}
\usepackage{bm}
\usepackage{fixltx2e}
% \usepackage[inline]{showlabels}

% biblio options
\bibliographystyle{elsarticle-harv}
\biboptions{authoryear}

@@

% macro fix for multiple asterisks for corr author
\makeatletter
\def\@author#1{\g@addto@macro\elsauthors{\normalsize%
    \def\baselinestretch{1}%
    \upshape\authorsep#1\unskip\textsuperscript{%
      \ifx\@fnmark\@empty\else\unskip\sep\@fnmark\let\sep=,\fi
      \ifx\@corref\@empty\else\unskip\sep\@corref\let\sep=,\fi
      }%
    \def\authorsep{\unskip,\space}%
    \global\let\@fnmark\@empty
    \global\let\@corref\@empty  %% Added
    \global\let\sep\@empty}%
    \@eadauthor={#1}
}
\makeatother

@@

% get the version from git log
<<echo = FALSE, cache = FALSE>>=
raw <- system('git log -1', intern = TRUE)
date <- raw[grep('^Date', raw)]
date <- gsub('^Date:\\s*', '', date)
commit <- raw[grep('^commit', raw)]
commit <- gsub('^commit ', '', commit)
commit <- paste0('\\href{https://github.com/fawda123/identifiability/commit/', commit, '}{', commit, '}')
vers <- paste0('Version: ', date, ', ', commit)
@

@@

\begin{frontmatter}

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{\Sexpr{vers}}

\date{}

\author{Marcus W. Beck\corref{mycorrespondingauthor}}
\address{USEPA National Health and Environmental Effects Research Laboratory, Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{beck.marcus@epa.gov}

\author{John C. Lehrter}
\address{Dauphin Island Sea Lab, University of South Alabama, Dauphin Island, AL 36528}
\ead{jlehrter@disl.org}

\author{Lisa L. Lowe}
\address{Lockheed Martin IS \& GS - Civil supporting the USEPA, Research Triangle Park, NC 27709}
\ead{lowe.lisa@epa.gov}

\begin{abstract}
\noindent Bio-geo-chemical models are useful tools in environmental sciences that can guide management and policy-making. Consequently, significant time and resources are spent developing these models in system-specific contexts. The optimization of model parameters to maximize precision, including transferability of these models to different systems, are fundamental concerns in the development and application of these tools. This study provides a context for understanding quantitative limitations of coupled hydrodynamic-ecological models by evaluating parameter sensitivity and identifability of a \ac{zerod} unit of a larger spatio-temporal model of hypoxia on the \acl{lcs} of Gulf of Mexico.  The analysis provides a contrast of numeric and ecological certainty in parameter subsets using a systematic framework to infer larger trends in dissolved oxygen dynamics over time, having implications for understanding factors that contribute to environmental conditions that are detrimental to aquatic resources.  In particular, we focus on issues of parameter identifiability using local sensitivity analyses to provide quantitative descriptions of numerical constraints on model precision.  The sensitivity of state variables differed considerably with parameter changes, although most variables were responsive to changes in parameters that influenced planktonic growth rates.  Variation in sensitivity had a direct correspendence with identifiability, such that only small subsets of the complete parameter set were characterized as having unique effects on the model output. As a result, we provide a set of parameter selection heuristics that can be used to identify parameters for model calibration that depend on relative sensitivity and ecological categories within the biogeochemical equations. Although these concerns have been expressed in the literature, they are rarely explicitly addressed or included in evaluations of water quality models.  In addition to immediate implications for regional models, we provide a framework for describing the effects of parameter uncertainty and identifiability that can be applied to similar models to better inform environmental management.
\end{abstract}

\begin{keyword}
\ac{cgem}, \ac{gom}, Hypoxia, Identifiability, Sensitivity
\end{keyword}

\end{frontmatter}

\acresetall
\linenumbers

@@

\bibliography{refs}
\clearpage

@@

% parameter exclusion temp
<<exclex, eval = T, fig.cap = 'Identifiability (as $\\gamma$, \\cref{gameq}) for \\ac{do} of organic matter parameters for subset combinations in \\cref{fig:identplo}.  Identifiability is evaluated for subsets that excluded and included the parameters at the top of each plot. Identifiability of including all eight parameters is in \\cref{fig:identplo}. Grey lines indicate potential thresholds at $\\gamma = 10, 15$ for maximum acceptable identifiability.', fig.height = 7.25, fig.width = 7.25, echo = F, cache = TRUE, out.width = "0.8\\textwidth">>=
data(cat_ident_all)

### 2016-12-16 ident_manu.tex
elsarticle class

\documentclass[letterpaper,12pt,oneside]{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

@@

\usepackage{alltt}
\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{indentfirst}
\usepackage{titlesec}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{outlines}
\usepackage{lineno}
\usepackage{array}
\usepackage{times}
\usepackage{cleveref}
\usepackage{acronym}
\usepackage[position=t]{subfig}
\usepackage{paralist}
\usepackage[noae]{Sweave}
\usepackage{natbib}
\usepackage{array}
\usepackage{pdflscape}
\usepackage{bm}
\usepackage{fixltx2e}
% \usepackage[inline]{showlabels}
\bibpunct{(}{)}{,}{a}{}{,}

% page margins and section title formatting
\linespread{1.5}
\setlength{\footskip}{0.5in}
\titleformat*{\section}{\Large\bf\em}
\titleformat*{\subsection}{\singlespace\large\bf}
\titleformat*{\subsubsection}{\singlespace\normalsize\bf\em}
\titlespacing{\section}{0in}{0in}{0in}
\titlespacing{\subsection}{0in}{0in}{0in}
\titlespacing{\subsubsection}{0in}{0in}{0in}

@@

\raggedbottom
\linenumbers
\raggedright
\urlstyle{same}
\setlength{\parindent}{0.5in}
\renewcommand\refname{References \vspace{12pt}}

\begin{singlespace}
\title{{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, John C. Lehrter$^2$, Lisa L. Lowe$^3$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Email: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}}}
	  \\\\{\textit {\normalsize $^2$Dauphin Island Sea Lab, University of South Alabama}}
  \\{\textit {\normalsize Dauphin Island, AL 36528}}
	\\{\textit {\normalsize Phone: 251-861-2141 ext. 7567}}
	\\{\textit {\normalsize Email: \href{mailto:jlehrter@disl.org}{jlehrter@disl.org}}}
	\\\\{\textit {\normalsize $^3$Lockheed Martin IS \& GS - Civil, supporting the USEPA}}
	\\{\textit {\normalsize Resarch Triangle Park, NC 27709}}
	\\{\textit {\normalsize Phone: 919-541-3985,}}
	\\{\textit {\normalsize Email: \href{mailto:lowe.lisa@epa.gov}{lowe.lisa@epa.gov}}}
  \vspace{1in} 
  \\ Version Date:   Wed Dec 14 07:59:53 2016 -0600
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

\begin{abstract}
\noindent Bio-geo-chemical models are useful tools in environmental sciences that can guide management and policy-making. Consequently, significant time and resources are spent developing these models in system-specific contexts. The optimization of model parameters to maximize precision, including transferability of these models to different systems, are fundamental concerns in the development and application of these tools. This study provides a context for understanding quantitative limitations of coupled hydrodynamic-ecological models by evaluating parameter sensitivity and identifability of a \ac{zerod} unit of a larger spatio-temporal model of hypoxia on the \acl{lcs} of Gulf of Mexico.  The analysis provides a contrast of numeric and ecological certainty in parameter subsets using a systematic framework to infer larger trends in dissolved oxygen dynamics over time, having implications for understanding factors that contribute to environmental conditions that are detrimental to aquatic resources.  In particular, we focus on issues of parameter identifiability using local sensitivity analyses to provide quantitative descriptions of numerical constraints on model precision.  The sensitivity of state variables differed considerably with parameter changes, although most variables were responsive to changes in parameters that influenced planktonic growth rates.  Variation in sensitivity had a direct correspendence with identifiability, such that only small subsets of the complete parameter set were characterized as having unique effects on the model output. As a result, we provide a set of parameter selection heuristics that can be used to identify parameters for model calibration that depend on relative sensitivity and ecological categories within the biogeochemical equations. Although these concerns have been expressed in the literature, they are rarely explicitly addressed or included in evaluations of water quality models.  In addition to immediate implications for regional models, we provide a framework for describing the effects of parameter uncertainty and identifiability that can be applied to similar models to better inform environmental management.
\end{abstract}
Keywords: \ac{cgem}, \ac{gom}, Hypoxia, Identifiability, Sensitivity
\acresetall

@@

{\centering \includegraphics[width=\textwidth]{figs/exclex-1}

>>>>

\documentclass[preprint]{elsarticle}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

@@

\usepackage{alltt}
\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
\usepackage{lineno}
\usepackage{cleveref}
\usepackage{acronym}
\usepackage{paralist}
\usepackage{bm}
\usepackage{fixltx2e}
% \usepackage[inline]{showlabels}

% biblio options
\bibliographystyle{elsarticle-harv}
\biboptions{authoryear}

@@

\begin{frontmatter}

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Thu Dec 15 15:47:19 2016 -0600, \href{https://github.com/fawda123/identifiability/commit/f47752c773b8b0d1c90abe57d41f57038201ea42}{f47752c773b8b0d1c90abe57d41f57038201ea42}}

\date{}

\author{Marcus W. Beck\corref{mycorrespondingauthor}}
\address{USEPA National Health and Environmental Effects Research Laboratory, Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{beck.marcus@epa.gov}

\author{John C. Lehrter}
\address{Dauphin Island Sea Lab, University of South Alabama, Dauphin Island, AL 36528}
\ead{jlehrter@disl.org}

\author{Lisa L. Lowe}
\address{Lockheed Martin IS \& GS - Civil supporting the USEPA, Research Triangle Park, NC 27709}
\ead{lowe.lisa@epa.gov}

\begin{abstract}
\noindent Bio-geo-chemical models are useful tools in environmental sciences that can guide management and policy-making. Consequently, significant time and resources are spent developing these models in system-specific contexts. The optimization of model parameters to maximize precision, including transferability of these models to different systems, are fundamental concerns in the development and application of these tools. This study provides a context for understanding quantitative limitations of coupled hydrodynamic-ecological models by evaluating parameter sensitivity and identifability of a \ac{zerod} unit of a larger spatio-temporal model of hypoxia on the \acl{lcs} of Gulf of Mexico.  The analysis provides a contrast of numeric and ecological certainty in parameter subsets using a systematic framework to infer larger trends in dissolved oxygen dynamics over time, having implications for understanding factors that contribute to environmental conditions that are detrimental to aquatic resources.  In particular, we focus on issues of parameter identifiability using local sensitivity analyses to provide quantitative descriptions of numerical constraints on model precision.  The sensitivity of state variables differed considerably with parameter changes, although most variables were responsive to changes in parameters that influenced planktonic growth rates.  Variation in sensitivity had a direct correspendence with identifiability, such that only small subsets of the complete parameter set were characterized as having unique effects on the model output. As a result, we provide a set of parameter selection heuristics that can be used to identify parameters for model calibration that depend on relative sensitivity and ecological categories within the biogeochemical equations. Although these concerns have been expressed in the literature, they are rarely explicitly addressed or included in evaluations of water quality models.  In addition to immediate implications for regional models, we provide a framework for describing the effects of parameter uncertainty and identifiability that can be applied to similar models to better inform environmental management.
\end{abstract}

\begin{keyword}
\ac{cgem}, \ac{gom}, Hypoxia, Identifiability, Sensitivity
\end{keyword}

\end{frontmatter}

\acresetall
\linenumbers

@@

{\centering \includegraphics[width=0.8\textwidth]{figs/exclex-1}
